Fixing Vulnerabilities Hinders Maintainability

### 2020-01-30 paper/icpc20.tex
changing example to CVE-2014-3488

@@

Software quality characteristics include, but are not limited to functional correctnes,

@@

However, there is still a considerable amount of zero-day vulnerabilities being

@@

flaw that allowed callers to see and hear others without them picking up the call.

@@

\Sof{After dropping Keep Your Codebase Small the example changed to a positive impact - the example needs to be changed} As an example, consider the refactoring of the Datagram Transport Layer Security

@@

\medskip
\setcounter{lstannotation}{0}
\begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
\\CVE-$2014$-$3506$ vulnerability},label={lst:vuln}]
+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+  if(max_len < (unsigned long)s->max_cert_list)
+     return s->max_cert_list;
+  return max_len;
+ }

  static int dtls1_reassemble_fragment(
     SSL *s, struct hm_header_st* msg_hdr, int *ok){

// [snip]

- unsigned long frag_len = msg_hdr->frag_len, max_len;
- if((msg_hdr->frag_off+frag_len) > msg_hdr->msg_len)
-    goto err;

- if(DTLS1_HM_HEADER_LENGTH +
-    	SSL3_RT_MAX_ENCRYPTED_LENGTH < s->max_cert_list)
-    max_len = s->max_cert_list;
- else
-    max_len = DTLS1_HM_HEADER_LENGTH +
-   		SSL3_RT_MAX_ENCRYPTED_LENGTH;

+ unsigned long frag_len = msg_hdr->frag_len;

- if((msg_hdr->frag_off+frag_len) > max_len)
+ if((msg_hdr->frag_off+frag_len) > msg_hdr->msg_len ||
+ msg_hdr->msg_len > dtls1_max_handshake_message_len(s))
     goto err; /*!\annotation{lst:func5}!*/

  memset(seq64be,0,sizeof(seq64be));
  seq64be[6] = (unsigned char) (msg_hdr->seq>>8);
  seq64be[7] = (unsigned char) msg_hdr->seq;
  item = pqueue_find(s->d1->buffered_messages, seq64be);
  if(item == NULL){
     frag = dtls1_hm_fragment_new(msg_hdr->msg_len, 1);

// [snip]

 static int dtls1_process_out_of_seq_message(
     SSL *s, struct hm_header_st* msg_hdr, int *ok){

// [snip]

+ if(frag_len > dtls1_max_handshake_message_len(s))
+    goto err; /*!\annotation{lst:func4}!*/

  frag = dtls1_hm_fragment_new(frag_len, 0);

@@

@@

This work presents an empirical study on the impact of $607$ security
refactorings on the maintainability of $94$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in previous work~\cite{Olivari:2018}. Security refactorings 
significantly affect the software maintainability of open-source software, i.e., 
developers hinder maintainability when fixing vulnerabilities. We found this evidence in
$40.2\%$ of the studied cases. Regarding regular commits, it seems that overall
maintainability increases ($50.16\%$). However, there is still a very
considerable amount of cases that harm maintainability ($41.35\%$).

>>>>

\lstdefinestyle{JavaStyle} {
    language=Java,
    backgroundcolor=\color{white},   % choose the background color
    basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
    breaklines=true,                 % automatic line breaking only at whitespace
    captionpos=b,                    % sets the caption-position to bottom
    commentstyle=\color{mygray}\bfseries,    % comment style
    escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
    keywordstyle=\color{blue},       % keyword style
    stringstyle=\color{mymauve},     % string literal style
    frame=single,
    numbers=left,
    stepnumber=1,
    xleftmargin=2em,
    escapeinside={/*!}{!*/},
    moredelim=**[l][\color{mygreen}]{+\ },
    moredelim=*[l][\color{red}]{-\ }
}

@@

Software quality characteristics include, but are not limited to functional correctness,

@@

However, there is still a considerable amount of vulnerabilities being

@@

flaw that allowed callers to see and hear others without them picking up the call.%\cite{cacheOut}.

@@

As an example, consider the refactoring of the Datagram Transport Layer Security

@@

\medskip
\setcounter{lstannotation}{0}
\begin{lstlisting}[style={JavaStyle}, caption={Fix provided by Netty developers to the
\\CVE-$2014$-$3488$ vulnerability},label={lst:vuln}]

// [snip]
   for (;;) {
+       final int outAppBufSize = engine.getSession().getApplicationBufferSize();
+       final ByteBuffer outAppBuf;
+       if (nioOutAppBuf.capacity() < outAppBufSize) {
+              outAppBuf = ByteBuffer.allocate(outAppBufSize);
+       } else {
+              outAppBuf = nioOutAppBuf;
+       }

         try {
-               result = engine.unwrap(nioInNetBuf, nioOutAppBuf);
-               result = engine.unwrap(nioInNetBuf, outAppBuf);
// [snip]
                 break;
         } finally {
-               nioOutAppBuf.flip();
+              outAppBuf.flip();

                nettyInNetBuf.readerIndex(nettyInNetBufStartOffset + nioInNetBuf.position() - nioInNetBufStartOffset);

-               if (nioOutAppBuf.hasRemaining()) {
+              if (outAppBuf.hasRemaining()) {
                       if (nettyOutAppBuf == null) {
                              ChannelBufferFactory factory = ctx.getChannel().getConfig().getBufferFactory();
                              nettyOutAppBuf = factory.getBuffer(initialNettyOutAppBufCapacity);
                       }
-                    nettyOutAppBuf.writeBytes(nioOutAppBuf);
+                    nettyOutAppBuf.writeBytes(outAppBuf);
                 }
-              nioOutAppBuf.clear();
+              outAppBuf.clear();
          }
     }

@@

%\medskip
%\setcounter{lstannotation}{0}
%\begin{lstlisting}[style={JavaStyle}, caption={Fix provided by OpenSSL developers to the
%\\CVE-$2014$-$3506$ vulnerability},label={lst:vuln}]
%+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
%+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
%+  if(max_len < (unsigned long)s->max_cert_list)
%+     return s->max_cert_list;
%+  return max_len;
%+ }
%
%  static int dtls1_reassemble_fragment(
%     SSL *s, struct hm_header_st* msg_hdr, int *ok){
%
%// [snip]
%
%- unsigned long frag_len = msg_hdr->frag_len, max_len;
%- if((msg_hdr->frag_off+frag_len) > msg_hdr->msg_len)
%-    goto err;
%
%- if(DTLS1_HM_HEADER_LENGTH +
%-    	SSL3_RT_MAX_ENCRYPTED_LENGTH < s->max_cert_list)
%-    max_len = s->max_cert_list;
%- else
%-    max_len = DTLS1_HM_HEADER_LENGTH +
%-   		SSL3_RT_MAX_ENCRYPTED_LENGTH;
%
%+ unsigned long frag_len = msg_hdr->frag_len;
%
%- if((msg_hdr->frag_off+frag_len) > max_len)
%+ if((msg_hdr->frag_off+frag_len) > msg_hdr->msg_len ||
%+ msg_hdr->msg_len > dtls1_max_handshake_message_len(s))
%     goto err; /*!\annotation{lst:func5}!*/
%
%  memset(seq64be,0,sizeof(seq64be));
%  seq64be[6] = (unsigned char) (msg_hdr->seq>>8);
%  seq64be[7] = (unsigned char) msg_hdr->seq;
%  item = pqueue_find(s->d1->buffered_messages, seq64be);
%  if(item == NULL){
%     frag = dtls1_hm_fragment_new(msg_hdr->msg_len, 1);
%
%// [snip]
%
% static int dtls1_process_out_of_seq_message(
%     SSL *s, struct hm_header_st* msg_hdr, int *ok){
%
%// [snip]
%
%+ if(frag_len > dtls1_max_handshake_message_len(s))
%+    goto err; /*!\annotation{lst:func4}!*/
%
%  frag = dtls1_hm_fragment_new(frag_len, 0);
%
%\end{lstlisting}

@@

This work presents an empirical study on the impact of $1037$ security
patches on the maintainability of $267$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in previous work~\cite{Olivari:2018}. Security patches 
significantly affect the software maintainability of open-source software, i.e., 
developers hinder maintainability when patching vulnerabilities. We found this evidence in
$41.76\%$ of the studied cases. Regarding regular commits, it seems that overall
maintainability increases (\Sof{Complete}). However, there is still a very
considerable amount of cases that harm maintainability (\Sof{Complete}).
